---
title: "Alycante_analysis"
author: "Marcelo Hurtado"
output: html_document
date: "2024-02-15"
---

Load functions
```{r}
source("../src/environment_set.R")
```

Load data
```{r}
counts.norm = read.csv("~/Downloads/Counts_TPM.csv", row.names=1)
traitData = read.csv("~/Downloads/Clinical_Alycante.csv", row.names=1)
deconv = read.delim("~/Downloads/Deconvolution_Alycante.txt", row.names=1)
```

Renaming cause columns names cannot start with numbers
```{r}
prefix <- rep("Sample_" , ncol(counts.norm))
colnames(counts.norm) <- paste0(prefix, substring(colnames(counts.norm), 2))
rownames(deconv) = paste0(prefix, rownames(deconv))
rownames(traitData) = paste0(prefix, rownames(traitData))
```

CellTFusion

1. Transcription factor inference
```{r}
tfs = compute.TFs.activity(counts.norm)
```

2. TFs network construction
```{r}
network = compute.WTCNA(tfs) 
#Association of TFs modules with clinical data
compute.metada.association(network[[1]], traitData, pval = 0.05, width = 10) 
```

3. Modules characterization
```{r}
compute.modules.enrichment(counts.norm, tfs, network[[1]], network[[2]])
```

4. Pathways activity inference
```{r}
pathways = compute.pathway.activity(counts.norm)
#Association of TFs modules with pathways
compute.modules.relationship(network[[1]], pathways, "Pathways_Progeny-TFs_Modules", width = 15)
```

5. Deconvolution analysis
```{r}
deconvolution = fix.names.GEMDeCan(deconv)
dt = compute.deconvolution.analysis(deconvolution, corr = 0.7)
#Association of TFs modules with deconvolution 
compute.modules.relationship(network[[1]], dt[[1]], "Deconvolution-TFs_Modules", width = 45)
```

6. Identify cell groups from significant deconvolution features across TFs modules
```{r}
features_sig = compute.modules.relationship(network[[1]], dt[[1]], return = T)
cell_dendrograms = identify.cell.groups(dt[[1]], network[[1]], features_sig, height = 20) 
```

7. Cell groups analysis
```{r}
res_anova_CHEMO = cell.groups.analysis(dt[[1]], traitData, "CHEMO", cell_dendrograms, cut.height = 0.7, stat.test = 'Anova', pval = 0.05, height = 20)
res_anova_CART = cell.groups.analysis(dt[[1]], traitData, "CART", cell_dendrograms, cut.height = 0.7, stat.test = 'Anova', pval = 0.05, height = 20)
```

Heatmap deconvolution
```{r}
ann_colors = HeatmapAnnotation(
  CHEMO = traitData$CHEMO,
  CART = traitData$CART,
  Time = traitData$time,
  Batch = traitData$Batch,
  
  col = list(CHEMO = c("POST_NR" = "green", "PRE_NR" = "red", "PRE_RESP" = "purple"),
             CART = c("CMR" = "yellow", "DEATH" = "blue", "NONE" = "black", "PMR" = "green", "PROGRESS" = "red"),
             Time = c("DIAG" = "black", "Rechute" = "purple"),
             Batch = c("1" = "green", "2" = "red")),
  annotation_legend_param = list(labels_gp = gpar(fontsize = 16), legend_width = unit(12, "cm"), 
                                 legend_heigh = unit(12, "cm"), title_gp = gpar(fontsize = 16))
    
)
 
ht1 = Heatmap(t(scale(dt[[1]])), top_annotation = ann_colors,
              border = T, name = "Deconvolution",
              clustering_distance_columns = "euclidean", clustering_method_rows = "ward.D2", clustering_method_columns = "ward.D2",
              column_dend_height = unit(5, "cm"), row_dend_width = unit(5, "cm"), 
              column_dend_reorder = T, row_dend_reorder = T,
              show_row_names = T, 
              show_heatmap_legend = T, 
              row_names_gp = gpar(fontsize = 15), 
              column_names_gp = gpar(fontsize =15), 
              width = unit(35, "cm"), height = unit(50, "cm"),
              heatmap_legend_param = list(labels_gp = gpar(fontsize = 12), legend_width = unit(12, "cm"), 
                                          legend_heigh = unit(12, "cm"), title_gp = gpar(fontsize = 12)))
  
pdf("Results/Deconvolution_after_subgrouping_annotated", width = 30, height = 28)
draw(ht1, show_heatmap_legend = T, heatmap_legend_side = "left", annotation_legend_side = 'left')
dev.off()
```

Save results
```{r}
save(tfs, network, pathways, dt, cell_dendrograms, res_anova_CART, res_anova_CHEMO, file = "results.RData")
```
