---
title: "EASIER_response"
output: html_document
date: "2025-01-23"
---

```{r, eval=TRUE}
BiocManager::install("easier", dependencies = TRUE)
library("easier")
```

```{r, message=FALSE}
suppressPackageStartupMessages({
  library("easierData")
  library("SummarizedExperiment")
})
```

## Compute hallmarks of immune response
```{r, eval=TRUE}
hallmarks_of_immune_response <- c("CYT", "Roh_IS", "chemokines", "Davoli_IS", "IFNy", "Ayers_expIS", "Tcell_inflamed", "RIR", "TLS")
immune_response_scores <- compute_scores_immune_response(RNA_tpm = counts.norm, 
                                                         selected_scores = hallmarks_of_immune_response)
head(immune_response_scores)
```

## Compute quantitative descriptors of the TME
```{r, eval=TRUE}
cell_fractions <- compute_cell_fractions(RNA_tpm = counts.norm)
head(cell_fractions)
```

By applying PROGENy [@HOLLAND2020194431; @Schubert2018] method to count data from RNA-seq, the activity of 14 signaling pathways can be inferred as in the chunk below. 
```{r, eval=TRUE}
pathway_activities <- compute_pathway_activity(RNA_counts = counts.norm,
                                               remove_sig_genes_immune_response = TRUE)
head(pathway_activities)
```

By applying DoRothEA [@Garcia-Alonso01082019] method to TPM data from RNA-seq, the activity of 118 TFs can be inferred as follows:
```{r, eval=TRUE}
tf_activities <- compute_TF_activity(RNA_tpm = counts.norm)
head(tf_activities[,1:5])
```

Using derived cancer-specific inter-cellular networks, the quantification of 867 ligand-receptor pairs can be done as in the chunk below. 
```{r, eval=TRUE}
lrpair_weights <- compute_LR_pairs(RNA_tpm = counts.norm,
                                   cancer_type = "pancan")
head(lrpair_weights[,1:5])
```

Using the ligand-receptor weights as input, 169 cell-cell interaction scores can be derived as in the chunk below. 
```{r, eval=TRUE}
ccpair_scores <- compute_CC_pairs(lrpairs = lrpair_weights, 
                                  cancer_type = "pancan")
head(ccpair_scores[,1:5])
```

Prediction
```{r, eval=TRUE}
opt_models = get_opt_models()
predictions <- predict_immune_response(pathways = pathway_activities,
                                       immunecells = cell_fractions,
                                       tfs = tf_activities,
                                       lrpairs = lrpair_weights,
                                       ccpairs = ccpair_scores,
                                       cancer_type = "LUAD", 
                                       verbose = TRUE)
```

## Evaluate easier predictions using patients' immunotherapy response

```{r, eval=TRUE}
response = as.character(coldata[,"R.NR"])
names(response) = rownames(coldata)
output_eval_with_resp <- assess_immune_response(predictions_immune_response = predictions,
                                                patient_response = response,
                                                RNA_tpm = counts.norm)
output_eval_with_resp
```

Heatmap to visualize predictions
```{r}
ann_colors <- HeatmapAnnotation(
  Response = coldata$R.NR,
  
  col = list(
    Response = c("R" = "green", "NR" = "red")
  ),
  
  annotation_name_gp = gpar(fontsize = 20),
  annotation_name_side = "right",
  gap = unit(5, "mm"),

  annotation_legend_param = list(labels_gp = gpar(fontsize = 20), legend_width = unit(15, "cm"),
                                 legend_side = "left", 
                                 title_gp = gpar(fontsize = 20))
  
)

# train_data_df = train_data %>%
#   dplyr::select(-target)

ht1 = Heatmap(t(scale(predictions$pathways)), border = T, top_annotation = ann_colors, 
              column_gap = unit(8, "mm"), name = "Deconvolution scores",
              clustering_method_rows = "ward.D2", 
              column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"), 
              column_dend_reorder = T, row_dend_reorder = F,
              show_row_names = T, 
              show_heatmap_legend = T, 
              row_names_gp = gpar(fontsize = 10), 
              column_names_gp = gpar(fontsize =10), 
              width = unit(40, "cm"), height = unit(40, "cm"),
              heatmap_legend_param = list(labels_gp = gpar(fontsize = 12), legend_width = unit(12, "cm"), 
                                          legend_heigh = unit(12, "cm"), title_gp = gpar(fontsize = 12)))

pdf(paste0("Results/Train32_", file_name), height = 20, width = 25)
draw(ht1, show_heatmap_legend = T, heatmap_legend_side = "left", annotation_legend_side = 'left')
dev.off()
```

Cell groups
```{r}
library(h2o)

lis_auc = c()

for(i in 1:2){
  set.seed(i) 
  
  clinical = coldata
  trait = "R.NR"
  trait.positive = "R"
  partition = 0.75
  
  # Do stratified partition 
  index = createDataPartition(clinical[,trait], times = 1, p = partition, list = FALSE) 
  
  # Train cohort
  traitData_train = clinical[index, ]
  counts.normalized_train = counts.norm[,index]
  
  # Test cohort
  traitData_test = clinical[-index, ]
  counts.normalized_test = counts.norm[,-index]
  
  #0.6714286 0.7857143 0.8285714 0.9285714 0.7428571 0.6571429 0.8428571 0.7928571 0.7071429 0.7571429
  #features_tme <- compute_scores_immune_response(RNA_tpm = counts.normalized_train,                          
  #                                                       selected_scores = hallmarks_of_immune_response)
  
  # features_tme <- compute_cell_fractions(RNA_tpm = counts.normalized_train)                  # 0.6857143 0.6285714
  # 
  # 
  # features_tme <- compute_pathway_activity(RNA_counts = counts.normalized_train,              # 0.5857143 0.5857143
  #                                                remove_sig_genes_immune_response = TRUE)
  # 
  #features_tme <- compute_TF_activity(RNA_tpm = counts.normalized_train) #0.7785714 0.8571429
  # 
  # lr <- compute_LR_pairs(RNA_tpm = counts.normalized_train, #0.7571429 0.7285714
  #                                    cancer_type = "pancan")
  # 
  # features_tme <- compute_CC_pairs(lrpairs = lr,
  #                                   cancer_type = "pancan") #0.5000000 0.4571429
  
  train_data = features_tme %>%
  data.frame() %>%
  mutate(Trait = traitData_train[,trait],
        target = as.factor(ifelse(Trait == trait.positive, 'yes', 'no'))) %>%
  dplyr::select(-Trait)
  
  train_data$target <- factor(train_data$target, levels = c("no", "yes"))  # Order, just in case to ensure positive class is not well defined
  
  features = setdiff(colnames(train_data), "target") 
  
  ####################### Testing set
  
  #testing_set <- compute_scores_immune_response(RNA_tpm = counts.normalized_test, 
  #                                                       selected_scores = hallmarks_of_immune_response)
  
  # testing_set <- compute_cell_fractions(RNA_tpm = counts.normalized_test)
  # 
  # 
  # testing_set <- compute_pathway_activity(RNA_counts = counts.normalized_test,
  #                                                remove_sig_genes_immune_response = TRUE)
  # 
  #testing_set <- compute_TF_activity(RNA_tpm = counts.normalized_test)
  # 
  lr <- compute_LR_pairs(RNA_tpm = counts.normalized_test,
                                     cancer_type = "pancan")

  testing_set <- compute_CC_pairs(lrpairs = lr,
                                    cancer_type = "pancan")
  
  target = traitData_test %>%
  mutate(target = ifelse(traitData_test[,trait] == trait.positive, "yes", "no")) %>%
  pull(target)
  
  target = factor(target, levels = c("no", "yes"))
    
  ########################################################### AutoML
  
  # Start the H2O cluster (locally)
  h2o.init()
  
  # Import a sample binary outcome train/test set into H2O
  train <- as.h2o(train_data)
  
  test <- testing_set %>%
  mutate(target = target) %>%
  as.h2o()
  
  # Identify predictors and response
  y <- "target"
  x <- setdiff(names(train), y)
  
  # For binary classification, response should be a factor
  train[, y] <- as.factor(train[, y])
  test[, y] <- as.factor(test[, y])
  
  # Run AutoML for 20 base models
  aml <- h2o.automl(x = x, y = y,
                  training_frame = train,
                  max_models = 20, exclude_algos = "DeepLearning",
                  seed = 123)
  
  best_model <- aml@leader
  
  # Generate predictions on the test set
  pred <- h2o.predict(best_model, test)
  
  # AUC from h2o
  perf <- h2o.performance(best_model, newdata = test)
  auc <- h2o.auc(perf)
  lis_auc = c(auc, lis_auc)
}

print(lis_auc)
```

```{r}
ann_colors <- HeatmapAnnotation(
  Response = train_data$target,
  
  col = list(
    Response = c("yes" = "green", "no" = "red")
  ),
  
  annotation_name_gp = gpar(fontsize = 20),
  annotation_name_side = "right",
  gap = unit(5, "mm"),

  annotation_legend_param = list(labels_gp = gpar(fontsize = 20), legend_width = unit(15, "cm"),
                                 legend_side = "left", 
                                 title_gp = gpar(fontsize = 20))
  
)

train_data_df = train_data %>%
  dplyr::select(-target)
ht1 = Heatmap(t(scale(train_data_df)), border = T, top_annotation = ann_colors, 
              column_gap = unit(8, "mm"), name = "Deconvolution scores",
              clustering_method_rows = "ward.D2", 
              column_dend_height = unit(2, "cm"), row_dend_width = unit(2, "cm"), 
              column_dend_reorder = T, row_dend_reorder = F,
              show_row_names = T, 
              show_heatmap_legend = T, 
              row_names_gp = gpar(fontsize = 10), 
              column_names_gp = gpar(fontsize =10), 
              width = unit(40, "cm"), height = unit(40, "cm"),
              heatmap_legend_param = list(labels_gp = gpar(fontsize = 12), legend_width = unit(12, "cm"), 
                                          legend_heigh = unit(12, "cm"), title_gp = gpar(fontsize = 12)))

pdf(paste0("Results/Train32_", file_name), height = 20, width = 25)
draw(ht1, show_heatmap_legend = T, heatmap_legend_side = "left", annotation_legend_side = 'left')
dev.off()
```
