---
title: "Template_script"
author: "Marcelo Hurtado"
date: "2024-10-22"
output: html_document
---

This script is used as a template for analyze your data with CellTFusion. The user needs to change the paths to specify the location of the following files: counts, traitData.

Set up environment
```{r setup, include=FALSE}
source("src/environment_set.R") #Load functions and packages
set.seed(123) #For reproducibility
```

Distribution
```{r}
#IMMUCAN
counts.norm.immucan = read.delim2("~/Downloads/all_cohorts_tpm_27112024.txt", row.names = 1)
rownames(counts.norm.immucan) <- str_replace(rownames(counts.norm.immucan), pattern = ".[0-9]+$", replacement = "")
counts.norm.immucan = convert_ensembl_symbol(counts.norm.immucan)
counts.norm.immucan = filter_coding_genes(counts.norm.immucan, annotation = "SYMBOL")
counts.norm.immucan.norm = ADImpute::NormalizeTPM(counts.norm.immucan)

#MIRACLE
counts.norm.miracle = read.csv("~/Downloads/salmon.merged.gene_tpm.tsv", sep = "\t", row.names = 1)
counts.norm.miracle$gene_name = NULL
```
```{r}
library(ggplot2)

gene_of_interest <- "MT1G"  
gene_expression <- as.numeric(counts.norm.immucan[gene_of_interest, ])  

ggplot(data.frame(Expression = gene_expression), aes(x = Expression)) +
  geom_density(fill = "blue", alpha = 0.3) +
  theme_minimal() +
  labs(title = "Gene expression distribution - IMMUCAN",
       x = "Expression value",
       y = "Density")

gene_expression <- as.numeric(counts.norm.immucan.norm[gene_of_interest, ])  

ggplot(data.frame(Expression = gene_expression), aes(x = Expression)) +
  geom_density(fill = "blue", alpha = 0.3) +
  theme_minimal() +
  labs(title = "Gene expression distribution - IMMUCAN normalize",
       x = "Expression value",
       y = "Density")
```

```{r}
library(ggplot2)

gene_of_interest <- "ENSG00000000003"  
gene_expression <- as.numeric(counts.norm.immucan[gene_of_interest, ])  

ggplot(data.frame(Expression = gene_expression), aes(x = Expression)) +
  geom_density(fill = "blue", alpha = 0.3) +
  theme_minimal() +
  labs(title = "Gene expression distribution - IMMUCAN",
       x = "Expression value",
       y = "Density")

gene_expression <- as.numeric(counts.norm.miracle[gene_of_interest, ])  

ggplot(data.frame(Expression = gene_expression), aes(x = Expression)) +
  geom_density(fill = "blue", alpha = 0.3) +
  theme_minimal() +
  labs(title = "Gene expression distribution - MIRACLE",
       x = "Expression value",
       y = "Density")
```

```{r}
deconv = compute.deconvolution(counts.norm.immucan.norm, normalized = F, methods = "Epidish", credentials.mail = "marcelo.hurtado@inserm.fr", credentials.token = "734212f6ad77fc4eea2bdb502792f294", file_name = "IMMUCAN")
```

Load data
```{r}
raw.counts = read.csv("../examples/Data/Mariathasan/Raw_files/Counts_Mariathasan_coding.csv", row.names = 1) 
traitData = read.csv("../examples/Data/Mariathasan/Raw_files/ClinicalData_Mariathasan.csv", row.names = 1)
```

We provided to the user an all-in-one function to compute CellTFusion, this will calculate and perform all intermediate analysis and finishing returning the cell groups scores. For this run the following:
```{r}
res = CellTFusion(raw.counts, cbsx.mail = "XXXXXXX", cbsx.token = "XXXXX", file_name = "Test")
```

If you prefer to do it step by step in order to see and analyze intermediate results, please refer to the following indications. 

CellTFusion needs as input cell-type deconvolution proportions and transcription factor activity. For this, we are gonna start by calculating these features.

0.1. Deconvolution

- Perform deconvolution using 7 methods and 10 signatures. It needs as input raw counts (genes as SYMBOL) and it will ask for your credentials to run CBSX
```{r}
deconv = compute.deconvolution(counts.norm, normalized = F, methods = c("Quantiseq", "Epidish", "DeconRNASeq"), credentials.mail = "marcelo.hurtado@inserm.fr", credentials.token = "734212f6ad77fc4eea2bdb502792f294", file_name = "Immucan")
```

If you don't want to run a specific method or used a specific signature, you can exclude them using the following:
```{r}
#The argument methods = () specify which methods you want to run 
deconv = compute.deconvolution(raw.counts, normalized = T, credentials.mail = "xxxx", credentials.token = "xxxxxx", methods = c("Quantiseq", "MCP", "XCell", "DWLS"), file_name = "Tutorial")
#The argument signatures_exclude specify which signature you want to exclude, this should be the name of the signature file
deconv = compute.deconvolution(raw.counts, normalized = T, credentials.mail = "xxxx", credentials.token = "xxxxxx", signatures_exclude = "BPRNACan", file_name = "Tutorial")
```

We have also implemented the option to use second-generation methods to deconvolve your data. These methods use as input single cell data to learn from it and deconvolve the bulk RNAseq. If you would like to use it, specify it in the parameters using the following.
```{r}
deconv = compute.deconvolution(raw.counts, normalized = T, credentials.mail = "xxxx", credentials.token = "xxxxxx", sc_deconv = T, sc_matrix = sc.object, cell_annotations = cell_labels, cell_samples = bath_ids, name_sc_signature = "Tutorial", file_name = "Tutorial")

#sc_deconv: Boolean value specifying if second-generation methods need to be run
#sc_matrix: Normalized gene expression matrix with genes as rows and cell as columns from the scRNAseq object
#cell_annotations: Vector corresponding with the cell annotations 
#cell_samples: Vector corresponding with the samples_ids of each cell
#name_sc_signature: Name you want to give to your signature
```

0.2. TFs inference: Needs normalized counts
```{r}
counts.norm = data.frame(ADImpute::NormalizeTPM(raw.counts, log = T)) 
tfs = compute.TFs.activity(counts.norm)
```

Once features have been calculated we can start analyzing our data using CellTFusion

1. TFs network construction
```{r}
network = compute.WTCNA(tfs, corr_mod = 0.9, clustering.method = "ward.D2", return = T) 
#Association of TFs modules with clinical data
compute.metada.association(network[[1]], traitData, pval = 0.05, width = 10) 
```

1.2. Modules characterization
```{r}
hub_tfs = identify_hub_TFs(t(tfs), network, MM_thresh = 0.8, degree_thresh = 0.9)
compute.modules.enrichment(counts.norm, hub_tfs)
```

2. Pathways activity inference
```{r}
pathways = compute.pathway.activity(counts.norm)
#Association of TFs modules with pathways
compute.modules.relationship(network[[1]], pathways, "Pathways_Progeny-TFs_Modules", width = 15)
```

3. Deconvolution analysis
```{r}
dt = compute.deconvolution.analysis(deconv, corr = 0.7, seed = 123) 
#Association of TFs modules with deconvolution 
compute.modules.relationship(network[[1]], dt[[1]], "Deconvolution-TFs_Modules", vertical = T, height = 30, width = 10, pval = 0.05)
```

4. Cell groups identification
```{r}
tfs.modules.clusters = compute.TF.network.classification(network, pathways, return = T)
corr_modules = compute.modules.relationship(network[[1]], dt[[1]], return = T, plot = F)
cell_dendrograms = identify.cell.groups(corr_modules, tfs.modules.clusters, height = 20, return = T)
```

5. Cell groups scores
```{r}
cell.groups = cell.groups.computation(dt[[1]], tfs.module.network = network, cell.dendrograms = cell_dendrograms) 
```

Analize CellTFusion results

Cell groups scores analysis to test association with a clinical trait (Anova or Fisher)
```{r}
res_fisher = cell.groups.fisher.test(cell.groups, traitData, "Response")
res_anova = cell.groups.anova.test(cell.groups, traitData, "Response")
```

Machine Learning models for predicting a clinical trait using the cell groups scores.
```{r}
# Run ML pipeline with one train-test split (set seed for reproducibility)
res_ml = compute.ML(raw.counts = raw.counts, normalized = T, clinical = traitData, trait = "Best.Confirmed.Overall.Response", 
                    trait.positive = "CR", partition = 0.8, metric = "Accuracy", stack = T, feature.selection = F, 
                    seed = 1234, doParallel = T,  workers = 2, file_name = "Test", return = T)

# Run ML pipeline with multiple train-test splits for consistency
compute.bootstrap.ML(raw.counts = raw.counts, normalized = T, clinical = traitData, trait = "Best.Confirmed.Overall.Response", 
                     trait.positive = "CR", partition = 0.8, metric = "Accuracy", iterations = 2, feature.selection = F, 
                     stack = T, workers = 2, file.name = "Test", return = F)
get_pooled_roc_curves("Test") #Get a boxplot with AUC scores across iterations

# Run ML using Leaving-one-dataset-out (LODO) approach
res_ml = compute.LODO.ML(raw.counts = raw.counts, normalized = T, clinical = traitData, trait = "Response", 
                         trait.positive = "R", trait.out = "Cohort", out = cohort, metric = "Accuracy", stack = T, 
                         feature.selection = F, doParallel = T, workers = 4, file_name = "Test", return = F)
```

