---
title: "Methylation deconvolution methods"
author: "Marcelo Hurtado"
date: "2026-02-21"
output: html_document
---

Load libraries
```{r}
source("~/Documents/multimethconv/R/methyl_deconvolution.R")
library(deconvMe)
```

Load example datasets
```{r}
methyl_set <- minfiData::MsetEx
ratio_set <- minfi::ratioConvert(methyl_set)
beta_matrix <- minfi::getBeta(ratio_set)
```

Compute deconvolution
```{r}
deconv = compute.deconvolution(methyl_set, doParallel = T, workers = 3, file_name = "Test")
```

Deconvolution methods

- EpiDISH (RPC, CP, SVR)
- Houseman
- MethylCC
- MethylResolver
- MethAtlas

EpiDISH
```{r}
result <- deconvMe::deconvolute(methyl_set = methyl_set, method = 'epidish', mode = "CBS")
result <- deconvMe::deconvolute(methyl_set = methyl_set, method = 'epidish', mode = "RPC")
result <- deconvMe::deconvolute(methyl_set = methyl_set, method = 'epidish', mode = "CP")
```

Houseman
```{r}
result <- deconvMe::deconvolute(methyl_set = methyl_set, method = 'houseman')
```

MethylCC
```{r}
result <- deconvMe::deconvolute(methyl_set = methyl_set, method = 'methylcc')
```

MethylResolver
```{r}
result <- deconvMe::deconvolute(methyl_set = methyl_set, method = 'methylresolver')
```

MethAtlas
```{r}
result <- deconvMe::deconvolute(methyl_set = methyl_set, method = 'methatlas')
```

Accessing signatures matrices in deconvME
```{r}
library(deconvMe)
library(dplyr)
library(tidyr)
library(ggplot2)
options(matrixStats.useNames.NA = "deprecated")

epidish_sig <- get_epidish_signature_matrix("blood")
houseman_sig <- get_houseman_signature_matrix()
methatlas_sig <- get_methatlas_signature_matrix()
methylresolver_sig <- get_methylresolver_signature_matrix()

write.table(epidish_sig, "signatures/epi-sig.txt", row.names = F)
write.table(houseman_sig, "signatures/houseman-sig.txt", row.names = F)
write.table(methatlas_sig, "signatures/methatlas-sig.txt", row.names = F)
write.table(methylresolver_sig, "signatures/methylresolver-sig.txt", row.names = F)
```

Cross-Signature Deconvolution

deconvMe supports cross-signature deconvolution, allowing you to use signature matrices from one method with the deconvolution algorithm of another method.

Supported Methods

Method	Cross-Signature Support
EpiDISH	✅ All methods
MethAtlas	✅ All methods
MethylResolver	✅ All methods
Houseman	❌ Not supported
MethylCC	❌ Not supported (no classical signature matrix)

```{r}
# EpiDISH with Houseman signature
result_epidish_houseman <- run_epidish(beta_matrix, reference = houseman_sig)

# MethAtlas with EpiDISH signature
result_methatlas_epidish <- run_methatlas(beta_matrix, reference = epidish_sig)

# MethylResolver with MethAtlas signature
result_mr_methatlas <- run_methylresolver(
  beta_matrix, 
  reference = methatlas_sig,
  alpha = 1
)
```

