
Load libraries
```{r}
#imports
library(readr)

#core methods
library(cosmosR)
library(MOFA2)
library(liana)
library(decoupleR)

#data manipulations
library(dplyr)
library(reshape2)
library(GSEABase)
library(tidyr)

#plotting
library(ggplot2)
library(ggfortify)
library(pheatmap)
library(gridExtra)
library(RColorBrewer)
# library(RCy3)
```

From omics data to MOFA input 
```{r}
mofa_input = read.csv("~/Documents/Factor_COSMOS/data/mofa/mofa_data.csv")
```

Run MOFA (don't run, results are already in the repository)
```{r}
for(i in c(5:15,20,length(unique(mofa_input$sample)))){
    wd <- getwd()
    write_MOFA_options_file(input_file = '/data/mofa/mofa_data.csv', output_file = paste0('/results/mofa/mofa_res_',i,'factor.hdf5'), factors = i, convergence_mode = "slow", likelihoods = c('gaussian','gaussian','gaussian'))
    system(paste0(paste('python3', wd, sep = " "), paste('/scripts/mofa/mofa2.py', "options_MOFA.csv")))
}
```

Inspect MOFA results
```{r}
# Load MOFA output
for(i in c(5:15,20,length(unique(mofa_input$sample)))){
  filepath <- paste0('~/Documents/Factor_COSMOS/results/mofa/mofa_res_',i,'factor.hdf5')
  model <- load_model(filepath)
  meta_data <- read_csv("~/Documents/Factor_COSMOS/data/metadata/RNA_metadata_cluster.csv")[,c(1,2)]
  colnames(meta_data) <- c("sample","cluster")
  mofa_metadata <- merge(samples_metadata(model), meta_data, by = "sample")
  names(mofa_metadata) <- c("sample","group","cluster")
  samples_metadata(model) <- mofa_metadata
  assign(paste0("model",i),model)
}
```

Plot factor values
```{r}
## Top 20 factor weights per view
grid.arrange(
  ### RNA
  plot_top_weights(model, 
                   view = "RNA",
                   factor = 4,
                   nfeatures = 10) +
    ggtitle("RNA"),
  ### Proteomics
  plot_top_weights(model, 
                   view = "proteo",
                   factor = 4,
                   nfeatures = 10) + 
    ggtitle("Proteomics"),
  ### Metabolomics
  plot_top_weights(model, 
                   view = "metab",
                   factor = 4,
                   nfeatures = 10) + 
    ggtitle("Metabolomics"), ncol =3
)
```

From MOFA output to TF activity, ligand-receptor activity

Extract the weights from our MOFA model and keep the weights from Factor 4 in the RNA view.
```{r}
weights <- get_weights(model, views = "all", factors = "all")
```

RNA weights 
```{r}
RNA <- data.frame(weights$RNA[,4]) 
row.names(RNA) <- gsub("_RNA","",row.names(RNA))
```

Metabolite weights
```{r}
metabs_all <- as.data.frame(weights$metab) 
row.names(metabs_all) <- gsub("_metab","",row.names(metabs_all))

#extract top metab weights
metabs_top <- metabs_all[order(apply(metabs_all,1,function(x){max(abs(x))}), decreasing = T)[1:10],]
```

Then we load the consensus networks from our databases, as well as the TF-targets regulons from collectri.
```{r}
# Load LIANA (receptor and ligand) consensus network
# ligrec_ressource <- distinct(liana::decomplexify(liana::select_resource("Consensus")[[1]]))
# save(ligrec_ressource, file = "support/ligrec_ressource.RData")

#process the LR interaction prior knowledge to make suitable input for decoupleR
load(file = "~/Documents/Factor_COSMOS/support/ligrec_ressource.RData")
ligrec_geneset <- format_LR_ressource(ligrec_ressource)

# Load Dorothea (TF) network
# dorothea_df <- decoupleR::get_collectri()
# save(dorothea_df, file = "support/dorothea_df.RData")
load(file = "~/Documents/Factor_COSMOS/support/dorothea_df.RData")
```

Calculate regulatory activities with decoupleR
```{r}
# Calculate regulatory activities from Receptor and Ligand network
ligrec_high_vs_low <- run_ulm(mat = as.matrix(RNA), network = ligrec_geneset, .source = set, .target = gene, minsize = 2) 
ligrec_high_vs_low <- ligrec_high_vs_low[ligrec_high_vs_low$statistic == "ulm",]
ligrec_high_vs_low_vector <- ligrec_high_vs_low$score
names(ligrec_high_vs_low_vector) <- ligrec_high_vs_low$source

# Calculate regulatory activities from TF network
TF_high_vs_low <- run_ulm(mat = as.matrix(RNA), network = dorothea_df, minsize = 10)
TF_high_vs_low <- TF_high_vs_low[TF_high_vs_low$statistic == "ulm",]
TF_high_vs_low_vector <- TF_high_vs_low$score
names(TF_high_vs_low_vector) <- TF_high_vs_low$source

## Save results
ligrec_TF_moon_inputs <- list("ligrec" = ligrec_high_vs_low_vector,
                              "TF" = TF_high_vs_low_vector)

```

Extracting network with COSMOS to formulate mechanistic hypotheses
```{r}
signaling_input <- ligrec_TF_moon_inputs$TF
ligrec_input <- ligrec_TF_moon_inputs$ligrec

RNA_input <- weights$RNA[,4] 
prot_input <- weights$proteo[,4]
metab_inputs <- weights$metab[,4]

#remove MOFA tags from feature names
names(RNA_input) <- gsub("_RNA","",names(RNA_input))
names(prot_input) <- gsub("_proteo","",names(prot_input))
```

MOON (meta footprint analysis)

In this part, we will run the network scoring to generate mechanistic hypotheses connecting ligands, receptors, TFs and metabolites, while pruning the network with information from both RNA and protein abundance to constrain the flux of signal consistent with the molecular data available.

In this approach, the network is compressed to avoid having redundant paths in the network leading to inflating the importance of some input measurements.

Processing of the PKN
```{r}
load("~/Documents/Factor_COSMOS/support/meta_network.RData")

### Remove unexpressed genes from the PKN
expressed_genes <- as.data.frame(read_csv("~/Documents/Factor_COSMOS/data/RNA/RNA_log2_FPKM_clean.csv"))
expressed_genes <- setNames(rep(1,length(expressed_genes[,1])), expressed_genes$Genes)
meta_network_filtered <- cosmosR:::filter_pkn_expressed_genes(names(expressed_genes), meta_pkn = meta_network)

##format metab inputs
metab_inputs <- as.numeric(scale(weights$metab[,4], center = F))
names(metab_inputs) <- row.names(weights$metab)

metab_to_HMDB <- as.data.frame(
  read_csv("~/Documents/Factor_COSMOS/data/metabolomic/MetaboliteToHMDB.csv"))

metab_to_HMDB <- metab_to_HMDB[metab_to_HMDB$common %in% names(metab_inputs),]
metab_inputs <- metab_inputs[metab_to_HMDB$common]
names(metab_inputs) <- metab_to_HMDB$HMDB

metab_inputs <- cosmosR::prepare_metab_inputs(metab_inputs, compartment_codes = c("m","c"))

#prepare upstream inputs
upstream_inputs <- c(rec_inputs) #the upstream input should be filtered for most significant
TF_inputs <- scale(ligrec_TF_moon_inputs$TF, center = F)
TF_inputs <- setNames(TF_inputs[,1], row.names(TF_inputs))
downstream_inputs <- c(metab_inputs, TF_inputs) #the downstream input should be complete

upstream_inputs_filtered <- cosmosR:::filter_input_nodes_not_in_pkn(upstream_inputs, meta_network_filtered)
downstream_inputs_filtered <- cosmosR:::filter_input_nodes_not_in_pkn(downstream_inputs, meta_network_filtered)

n_steps <- 6

# in this step we prune the network to keep only the relevant part between upstream and downstream nodes
meta_network_filtered <- cosmosR:::keep_controllable_neighbours(meta_network_filtered
                                                       , n_steps, 
                                                       names(upstream_inputs_filtered))


downstream_inputs_filtered <- cosmosR:::filter_input_nodes_not_in_pkn(downstream_inputs_filtered, meta_network_filtered)


meta_network_filtered <- cosmosR:::keep_observable_neighbours(meta_network_filtered, n_steps, names(downstream_inputs_filtered))


upstream_inputs_filtered <- cosmosR:::filter_input_nodes_not_in_pkn(upstream_inputs_filtered, meta_network_filtered)


#write_csv(meta_network_filtered, file = "results/cosmos/moon/meta_network_filtered.csv")

#compress the network to avoid redundant master controllers
meta_network_compressed_list <- compress_same_children(meta_network_filtered, sig_input = upstream_inputs_filtered, metab_input = downstream_inputs_filtered)

meta_network_compressed <- meta_network_compressed_list$compressed_network

meta_network_compressed <- meta_network_cleanup(meta_network_compressed)

load(file = "support/dorothea_df.RData")

cosmosR::moon()
```