---
title: "Omnideconv"
output: html_document
date: "2024-10-16"
---

```{r}
library(Seurat)
library(SeuratDisk)
library(patchwork)
library(ggplot2)
library(dplyr)
library(omnideconv)
set.seed(123)
```

```{r}
counts = read.csv("~/Downloads/RawFiles/Counts_Vanderbilt.csv", row.names = 1) 
sc_rnaseq = LoadH5Seurat("~/all_samples.h5seurat",  assays = "RNA")
sc_rnaseq <- subset(sc_rnaseq, subset = n_genes > 200 & total_counts < 2500 & pct_counts_mt < 5)
sc_rnaseq <- sc_rnaseq[1:50, sample(100, 50)]
```

```{r}
library(stringr)
rownames(counts) <- str_replace(rownames(counts),
                        pattern = ".[0-9]+$",
                        replacement = "")
```

Extract only coding genes
```{r}
library(biomaRt)

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

gene_ids <- rownames(counts)

# Get gene type annotations
annotations <- getBM(attributes = c("ensembl_gene_id", "gene_biotype"), 
                     filters = "ensembl_gene_id", 
                     values = gene_ids, 
                     mart = mart)

# Filter annotations to keep only coding genes
coding_genes <- annotations %>%
  filter(gene_biotype == "protein_coding") %>%
  pull(ensembl_gene_id)

counts = counts[rownames(counts) %in% coding_genes,]

# Change to gene SYMBOL
library(org.Hs.eg.db)
library(plotly)
library(tibble)

entrz <- AnnotationDbi::select(org.Hs.eg.db, keys = rownames(counts), columns = "SYMBOL", keytype = "ENSEMBL") %>%
  distinct(SYMBOL, .keep_all=T) %>%
  filter(ENSEMBL %in% rownames(counts)) %>%
  group_by(ENSEMBL) %>%
  slice(1) %>%
  ungroup()

counts = counts %>%
  rownames_to_column("Genes") %>%
  filter(Genes %in% entrz$ENSEMBL) %>%
  mutate(Symbol = entrz$SYMBOL) %>%
  na.omit() %>%
  remove_rownames() %>%
  column_to_rownames("Symbol") %>%
  dplyr::select(-Genes)

# Get gene type annotations
annotations_symbol <- getBM(attributes = c("hgnc_symbol", "gene_biotype"),
                     filters = "hgnc_symbol",
                     values = rownames(counts),
                     mart = mart)

# Filter annotations to keep only coding genes
coding_genes <- annotations_symbol %>%
  filter(gene_biotype == "protein_coding") %>%
  pull(hgnc_symbol)

counts = counts[rownames(counts) %in% coding_genes,]
norm_counts = ADImpute::NormalizeTPM(counts, log=T) 
```

```{r}
sc_rnaseq = sc_rnaseq[rownames(sc_rnaseq)%in%coding_genes,]
```

```{r}
single_cell_object = as.matrix(sc_rnaseq@assays$RNA@data)
single_cell_object_scale = as.matrix(sc_rnaseq@assays$RNA@scale.data)
cell_type_annotations = sc_rnaseq@meta.data$annotated_ct
batch_ids = sc_rnaseq@meta.data$sample

rm(sc_rnaseq)
gc()
```

Upload signature
```{r}
sig_file = read.delim('~/Documents/CellTFusion/user_projects/src/signatures/CBSX-Melanoma-scRNAseq.txt', row.names = 1)
genes = rownames(sig_file)

sig_file <- sig_file %>%
  apply(., 2, as.numeric) %>%
  data.frame() %>%
  mutate("Genes" = genes) %>%
  column_to_rownames("Genes") %>%
  as.matrix()
```

AutoGeneS: Need single cell object
```{r}
#signature <- omnideconv::build_model(single_cell_object_scale, as.character(cell_type_annotations), method = "autogenes") #takes too much time
deconvolution = deconvolute_autogenes(bulk_gene_expression = norm_counts, single_cell_object = single_cell_object_scale,
                                      cell_type_annotations = as.character(cell_type_annotations), verbose = T)
head(deconvolution$proportions)
```

BayesPrism: Need single cell object
```{r}
signature <- omnideconv::build_model(single_cell_object, as.character(cell_type_annotations), method = "bayesprism") #does not return signature
deconvolution = deconvolute_bayesprism(bulk_gene_expression = counts, single_cell_object = single_cell_object,
                                       cell_type_annotations = cell_type_annotations, n_cores = 8)
```

Bisque: Need single cell object
```{r}
signature <- omnideconv::build_model(single_cell_object, as.character(cell_type_annotations), method = "bisque") #does not return signature
deconvolution = deconvolute_bisque(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object,
                                   cell_type_annotations = cell_type_annotations, batch_ids = batch_ids, verbose = T)
```

Bseq-Sc: You can input the matrix
```{r}
deconvolution = deconvolute_bseqsc(bulk_gene_expression = counts, signature = sig_file, 
                                   log = T, verbose = T)
#Need CIBERSORT source R code
```

CDseqR: Need single cell object
```{r}
signature <- omnideconv::build_model(single_cell_object, as.character(cell_type_annotations), method = "cdseq") #does not return signature
deconvolution = deconvolute_cdseq(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object, no_cores = 8,
                                  cell_type_annotations = cell_type_annotations, batch_ids = batch_ids, verbose = T) #errpr
```

CIBERSORTx: You can input the matrix
```{r}
deconvolution = deconvolute_cibersortx(bulk_gene_expression = counts, signature = sig_file)
```

CPM: Need single cell object
```{r}
signature <- omnideconv::build_model(single_cell_object, as.character(cell_type_annotations), method = "cpm") #does not return signature
deconvolution = deconvolute_cpm(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object,
                                cell_type_annotations = cell_type_annotations, no_cores = 8, verbose = T)
```

DWLS: You can input the matrix
```{r}
deconvolution = deconvolute_dwls(bulk_gene_expression = counts, signature = sig_file, 
                                 dwls_submethod = "DampenedWLS", verbose = T)
```

MOMF: Need single cell
```{r}
### compute the cell type specific expression level as reference
# signature = cell type specific expression levels reference used by MOMF --> priorU <- momf.computeRef(sc_counts, sc_cell_type)
signature <- omnideconv::build_model_momf(single_cell_object, as.character(cell_type_annotations), counts) #return signature
deconvolution = deconvolute_momf(bulk_gene_expression = counts, single_cell_object = single_cell_object, signature = signature, method = "KL", verbose = T)
```

MuSiC: Need single cell
```{r}
signature <- omnideconv::build_model(single_cell_object, as.character(cell_type_annotations), method = "music") #does not return signature
deconvolution = deconvolute_music(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object,
                                  cell_type_annotations = cell_type_annotations, batch_ids = batch_ids, verbose = T)
```

Scaden: Need single cell
```{r}
signature <- omnideconv::build_model(bulk_gene_expression = counts, single_cell_object, as.character(cell_type_annotations), method = "scaden") #does not return signature
deconvolution = deconvolute_scaden(bulk_gene_expression = counts, signature = '~/Documents/CellTFusion/user_projects/src/signatures/', verbose = T)
```

SCDC: Need single cell
```{r}
deconvolution = deconvolute_scdc(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object, 
                                  cell_type_annotations = cell_type_annotations, batch_ids = batch_ids, verbose = T)
```

Testing methods 
```{r}
sc_object = single_cell_object
bulk_counts = norm_counts
cell_annotations = cell_type_annotations
samples_ids = batch_ids
```

```{r}
message("Running second generation cell-type deconvolution methods using scRNAseq\n")

cat("\nRunning AutogeneS...............................................................\n")
dwls = deconvolute_autogenes(bulk_gene_expression = bulk_counts, single_cell_object = sc_object,
                             cell_type_annotations = as.character(cell_annotations), verbose = T)$proportions
cat("\nRunning BayesPrism...............................................................\n")
bayesprism = deconvolute_bayesprism(bulk_gene_expression = bulk_counts, single_cell_object = sc_object,
                              cell_type_annotations = cell_type_annotations, n_cores = 8) #raw counts
cat("\nRunning Bisque...............................................................\n")
bisque = deconvolute_bisque(bulk_gene_expression = as.matrix(bulk_counts), single_cell_object = sc_object,
                                cell_type_annotations = cell_type_annotations, batch_ids = samples_ids, verbose = T) #raw counts
# cat("\nRunning BSeq-sc...............................................................\n") #Need CIBERSORT source code
# bseqsc = deconvolute_bisque(bulk_gene_expression = as.matrix(bulk_counts), single_cell_object = sc_object,
#                             cell_type_annotations = cell_type_annotations, batch_ids = samples_ids, verbose = T)
cat("\nRunning CDSeq...............................................................\n")
cdseq = deconvolute_cdseq(bulk_gene_expression = as.matrix(bulk_counts), single_cell_object = sc_object, no_cores = 8, #raw counts
                           cell_type_annotations = cell_type_annotations, batch_ids = samples_ids, verbose = T)$cdseq_prop_merged_byCorr
cat("\nRunning CibersortX...............................................................\n")
set_cibersortx_credentials("marcelo.hurtado@inserm.fr", "734212f6ad77fc4eea2bdb502792f294")
model_cbsx <- omnideconv::build_model(sc_object, as.character(cell_type_annotations), batch_ids = samples_ids, method = "cibersortx") 
cibersortx = deconvolute_cibersortx(bulk_gene_expression = as.matrix(bulk_counts), signature = model_cbsx, verbose = T)
cat("\nRunning CPM...............................................................\n")
cpm = deconvolute_cpm(bulk_gene_expression = as.matrix(bulk_counts), single_cell_object = sc_object, no_cores = 8, #raw counts
                             cell_type_annotations = cell_type_annotations, verbose = T)
cat("\nRunning DWLS...............................................................\n")
model_dwls <- omnideconv::build_model(sc_object, as.character(cell_type_annotations), batch_ids = samples_ids, method = "dwls") 
deconvolution_dwls = deconvolute_dwls(bulk_gene_expression = bulk_counts, signature = model_dwls, dwls_submethod = "DampenedWLS", verbose = T)
cat("\nRunning MOMF...............................................................\n")
model_momf <- omnideconv::build_model(bulk_gene_expression = bulk_counts, sc_object, as.character(cell_type_annotations), batch_ids = samples_ids, method = "momf") #raw counts 
momf = deconvolute_momf(bulk_gene_expression = as.matrix(bulk_counts), single_cell_object = sc_object,
                              signature = model_momf, method = "KL", verbose = T)
cat("\nRunning MuSiC...............................................................\n")
music = deconvolute_music(bulk_gene_expression = as.matrix(bulk_counts), single_cell_object = sc_object,
                             cell_type_annotations = cell_type_annotations,  batch_ids = samples_ids, verbose = T)
cat("\nRunning Scaden...............................................................\n")
model_scaden <- omnideconv::build_model(bulk_gene_expression = bulk_counts, sc_object, as.character(cell_type_annotations), batch_ids = samples_ids, method = "scaden") 
scaden = deconvolute_scaden(bulk_gene_expression = bulk_counts, signature = model_scaden, verbose = T)
cat("\nRunning SCDC...............................................................\n")
scdc = deconvolute_scdc(bulk_gene_expression = as.matrix(bulk_counts), single_cell_object = sc_object, 
                        cell_type_annotations = cell_type_annotations, batch_ids = samples_ids, verbose = T)
```

