---
title: "Omnideconv"
output: html_document
date: "2024-10-16"
---

```{r}
library(Seurat)
library(SeuratDisk)
library(patchwork)
library(ggplot2)
library(dplyr)
set.seed(123)
```

```{r}
counts = read.csv("~/Documents/LungPredict1_paper/RawFiles/Counts_Vanderbilt.csv", row.names = 1) 
sc_rnaseq = LoadH5Seurat("~/all_samples.h5seurat",  assays = "RNA")
sc_rnaseq <- subset(sc_rnaseq, subset = n_genes > 200 & total_counts < 2500 & pct_counts_mt < 5)
sc_rnaseq <- sc_rnaseq[, 1:100]
```

```{r}
library(stringr)
rownames(counts) <- str_replace(rownames(counts),
                        pattern = ".[0-9]+$",
                        replacement = "")
```

Extract only coding genes
```{r}
library(biomaRt)

mart <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")

gene_ids <- rownames(counts)

# Get gene type annotations
annotations <- getBM(attributes = c("ensembl_gene_id", "gene_biotype"), 
                     filters = "ensembl_gene_id", 
                     values = gene_ids, 
                     mart = mart)

# Filter annotations to keep only coding genes
coding_genes <- annotations %>%
  filter(gene_biotype == "protein_coding") %>%
  pull(ensembl_gene_id)

counts = counts[rownames(counts) %in% coding_genes,]

# Change to gene SYMBOL
library(org.Hs.eg.db)
library(plotly)
library(tibble)

entrz <- AnnotationDbi::select(org.Hs.eg.db, keys = rownames(counts), columns = "SYMBOL", keytype = "ENSEMBL") %>%
  distinct(SYMBOL, .keep_all=T) %>%
  filter(ENSEMBL %in% rownames(counts)) %>%
  group_by(ENSEMBL) %>%
  slice(1) %>%
  ungroup()

counts = counts %>%
  rownames_to_column("Genes") %>%
  filter(Genes %in% entrz$ENSEMBL) %>%
  mutate(Symbol = entrz$SYMBOL) %>%
  na.omit() %>%
  remove_rownames() %>%
  column_to_rownames("Symbol") %>%
  dplyr::select(-Genes)

# Get gene type annotations
annotations_symbol <- getBM(attributes = c("hgnc_symbol", "gene_biotype"),
                     filters = "hgnc_symbol",
                     values = rownames(counts),
                     mart = mart)

# Filter annotations to keep only coding genes
coding_genes <- annotations_symbol %>%
  filter(gene_biotype == "protein_coding") %>%
  pull(hgnc_symbol)

counts = counts[rownames(counts) %in% coding_genes,]
```

```{r}
sc_rnaseq = sc_rnaseq[rownames(sc_rnaseq)%in%coding_genes,]
```

```{r}
single_cell_object = as.matrix(sc_rnaseq@assays$RNA@data)
cell_type_annotations = sc_rnaseq@meta.data$annotated_ct
batch_ids = sc_rnaseq@meta.data$sample

rm(sc_rnaseq)
gc()
```

Build model
```{r}
library(omnideconv)
signature <- omnideconv::build_model(single_cell_object, cell_type_annotations,
                                     batch_ids, method = "dwls", dwls_method = 'mast_optimized')
```

Upload signature
```{r}
sig_file = read.delim2('~/Documents/CellTFusion/user_projects/src/signatures/CBSX-Melanoma-scRNAseq.txt')
genes = sig_file$NAME

sig_file <- sig_file %>%
  column_to_rownames("NAME") %>%
  apply(., 2, as.numeric) %>%
  data.frame() %>%
  mutate("Genes" = genes) %>%
  column_to_rownames("Genes") %>%
  as.matrix()

```

AutoGeneS: Need single cell object
```{r}
deconvolution = deconvolute_autogenes(bulk_gene_expression = counts, single_cell_object = single_cell_object,
                                      cell_type_annotations = cell_type_annotations, verbose = T)
```

BayesPrism: Need single cell object
```{r}
deconvolution = deconvolute_bayesprism(bulk_gene_expression = counts, single_cell_object = single_cell_object,
                                       cell_type_annotations = cell_type_annotations, n_cores = 12)
```

Bisque: Need single cell object
```{r}
deconvolution = deconvolute_bisque(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object,
                                   cell_type_annotations = cell_type_annotations, batch_ids = batch_ids, verbose = T)
```

Bseq-Sc: You can input the matrix
```{r}
deconvolution = deconvolute_bseqsc(bulk_gene_expression = counts, signature = sig_file, 
                                   log = T, verbose = T)
```

CDseqR: Need single cell object
```{r}
deconvolution = deconvolute_cdseq(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object,
                                  cell_type_annotations = cell_type_annotations, batch_ids = batch_ids, verbose = T)
```

CIBERSORTx: You can input the matrix
```{r}
deconvolution = deconvolute_cibersortx(bulk_gene_expression = counts, signature = sig_file)
```

CPM: Need single cell object
```{r}
deconvolution = deconvolute_cpm(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object,
                                cell_type_annotations = cell_type_annotations, no_cores = 12, verbose = T)
```

DWLS: You can input the matrix
```{r}
deconvolution = deconvolute_dwls(bulk_gene_expression = counts, model = sig_file, 
                                 dwls_submethod = "DampenedWLS", verbose = T)
```

MOMF: You can input the matrix
```{r}
deconvolution = deconvolute_momf(bulk_gene_expression = counts, signature = sig_file, 
                                 method = "KL", verbose = T)
```

MuSiC: Need single cell
```{r}
deconvolution = deconvolute_music(bulk_gene_expression = as.matrix(counts), single_cell_object = single_cell_object,
                                  cell_type_annotations = cell_type_annotations, batch_ids = batch_ids, verbose = T)
```

Scaden: You can input the matrix
```{r}
deconvolution = deconvolute_scaden(bulk_gene_expression = counts, signature = sig_file, verbose = T)
```

SCDC: You can input the matrix
```{r}
deconvolution = deconvolute_scdc(bulk_gene_expression = counts, signature = sig_file, verbose = T)
```

