---
title: "PHiC_CpGs_analysis"
author: "Marcelo Hurtado"
date: "2026-02-22"
output: html_document
---

This script integrates DNA methylation data with PCHi-C interactions to identify genes with methylated promoters and genes whose promoters interact with methylated distal regions, generating gene sets for adding into RNAseq signatures.

Load libraries
```{r}
library(chaser)
library(dplyr)
source("functions.R")
```

Load EPIC manifest to map CpGs with chromosome regions
```{r}
load("../output/Manifest_EPIC_CpGs_chr.RData")
```

Load differential CpGs
```{r}
load("../data/DNAMethSignature/DNAMethSignature/all_CPGs_for_pairwise.RData")
list_cpg = FitList1[["Activated_Macrophage-B_cell"]]$ID
## Subset to only differential CpGs
cpg = manifest %>%
  dplyr::filter(CPGs %in% list_cpg)
```

Prepare CpG feature table from PCHi-C network
```{r}
cpgsel <- data.frame(
  chr   = cpg[,2],
  start = cpg[,3],
  end   = cpg[,4],
  value = 1 # Value = 1 means this genomic region contains a CpG belonging to the methylation signature.
)

# Why cpgsel is needed?

# Chaser:
# - Overlaps your CpG intervals with PCHi-C fragments
# - For each fragment, sums the value
# - Stores that number inside the network object
# 
# So after loading features:
# - If a fragment overlaps 3 CpGs → it gets value = 3
# - If it overlaps 0 CpGs → it gets value = 0
```

Builds a promoter–enhancer interaction network and annotates it with methylation information.
```{r}
load("../data/pchic.RData") #PCHi-C data: each row corresponds to one chromatin interaction: |bait(promoter)|other end (distal fragment)|

# Add chr prefix
pchic[,1] <- paste0('chr', pchic[,1]) #formatting consistency
pchic[,6] <- paste0('chr', pchic[,6]) #formatting consistency

bp_forchaser <- pchic[, c(1,2,3,6,7,8)] # Promoter coordinates  →  Distal coordinates

# Build chromatin network
netbp <- chaser::make_chromnet(bp_forchaser) # Each unique genomic fragment becomes a node / Each interaction becomes an edge
netbp <- chaser::load_features(netbp, cpgsel, type = 'features_table', missingv = 0) #Add methylation features to network

# What happens in load_features():
# - cpgsel contains CpG genomic coordinates.
# - chaser overlaps CpGs with PCHi-C fragments.
# For each fragment:
# - If overlaps CpG → value added
# - If no overlap → value = 0
# So now every node has a feature:
# - Fragment A → 2 CpGs
# - Fragment B → 0 CpGs
# - Fragment C → 1 CpG
```

Identify CpG-Containing Fragments
```{r}
netbpfeat <- chaser::export(netbp) #matrix with rows as fragments in the PCHi-C network (nodes) and value of CpGs overlapping that fragment
cpgnodes  <- rownames(netbpfeat)[netbpfeat > 0] #select CpG-containing nodes
```

Identify Promoters and Other Ends
```{r}
edges  <- chaser::export(netbp, "edges") #data frame of all interactions in the chromatin network

baits  <- unique(edges$node_from) #all fragments that act as promoters (nodes that initiate interactions)
oes    <- setdiff(unique(edges$node_to), baits) #all fragments that act as Other Ends (distal regions) excluding promoters

promcpg <- intersect(cpgnodes, baits) #promoters that contain at least one CpG
oecpg   <- intersect(cpgnodes, oes) #distal fragments that contain at least one CpG

### Summary:
# promcpg → directly methylated promoters
# oecpg → distal methylated fragments that may regulate promoters via chromatin looping
```

Promoter fragments and associated genes
```{r}
# Combines chromosome, start, and end coordinates into a single string.
pchic$fragmentID <- paste0(
  pchic$baitChr, ":",
  pchic$baitStart, "-",
  pchic$baitEnd
)

# Keep unique promoter fragments
bait_annotation <- unique( #unique() ensures that each fragmentID appears only once, even if it interacts with multiple distal fragments
  pchic[, c("fragmentID", "baitName")] # "baitName" contains gene names linked to each promoter fragment (multiple genes separated by ;)
) 

# Set rownames as fragment IDs
rownames(bait_annotation) <- bait_annotation$fragmentID
```

Genes with Methylated Promoters
```{r}
# Extract gene symbol
prom_genes_raw <- bait_annotation[promcpg, "baitName"]

# Split multiple IDs and clean
prom_genes <- unique(unlist(strsplit(prom_genes_raw, ";")))

write.table(prom_genes, '../output/promoter_CpGs_SYMBOL.txt', quote = FALSE, row.names = FALSE, col.names = FALSE)
```

Genes Whose Promoters Interact with Methylated Distal Regions
```{r}
# Convert to igraph
G <- igraph::graph_from_data_frame(
  edges[, c("node_from", "node_to")],
  directed = FALSE
)

# Find promoters interacting with methylated distal fragments
neighs <- unique(unlist(lapply(oecpg, function(x) { #oecpg = distal fragments that contain methylated CpGs.
  neighbors(G, x)$name #for each methylated distal fragment, get all connected promoter fragments
})))

# neighs = all promoter fragment IDs that interact with methylated distal fragments.

# Extract gene symbol
neigh_genes_raw <- bait_annotation[neighs, "baitName"]

# Split multiple IDs and clean
neigh_genes <- unique(unlist(strsplit(neigh_genes_raw, ";")))

write.table(neigh_genes, '../output/oe_CpGs_SYMBOL.txt', quote = FALSE, row.names = FALSE, col.names = FALSE)
```

Load CpGs
```{r}
prom_genes = read.table("../output/promoter_CpGs_SYMBOL.txt")$V1
neigh_genes = read.table("../output/oe_CpGs_SYMBOL.txt")$V1
```

Merge methylation-informed genes into RNA signature
```{r}
rna_signature = read.delim2("~/Documents/multideconv/inst/signatures/BPRNACan.txt")
rna_genes = rna_signature$Genes

# Identify genes overlapping with RNA signature
overlap_prom <- intersect(prom_genes, rna_genes)   # promoter-methylated genes already in RNA signature
overlap_neigh <- intersect(neigh_genes, rna_genes)    # distal-interacting genes already in RNA signature

# Create expanded RNA signature:
# - Combine original RNA signature with genes from methylation signature not already in RNA signature
expanded_signature <- unique(c(rna_genes, prom_genes, neigh_genes))
```

Expand RNA signature with CpG signature
```{r}
source("functions.R")
load("../output/CpG2genes.RData")
load("../data/BLUEPRINT/BP_WGBS.RData")
x = read.delim("~/Documents/multideconv/inst/signatures/BPRNACan.txt", row.names = 1)
x_expanded = compute_CpG_extension(x, Final.Tumor.sig, cpg_gene_map)
```

Add original TCGA data
```{r}
load("../data/TCGA_RNAseq_cancer/")
```

