---
title: "File"
output: html_document
date: "2024-12-03"
---

Input
```{r}
reference = readRDS("~/Documents/starting_kit_phase1/data/reference_pdac.rds")
mixes = readRDS("~/Documents/starting_kit_phase1/data/mixes1_insilicodirichletEMFA_pdac.rds")
```

Single cell object
```{r}
library(Seurat)
sc = list()
for (i in 1:length(reference$ref_scRNA)) {
  sc[[i]] = CreateSeuratObject(as.matrix(reference$ref_scRNA[[i]]$counts), metadata = reference$ref_scRNA[[i]]$metadata)
  sc[[i]]@meta.data$cell.annot = reference$ref_scRNA[[i]]$metadata$cell_type
  Idents(sc[[i]]) = sc[[i]]@meta.data$cell.annot
}
```

Merge scRNAseq
```{r}
library(SeuratData)
datasets <- c("Peng", "Baron", "Raghavan")
seuratobject <- merge(sc[[1]], y = c(sc[[2]], sc[[3]]), add.cell.ids = datasets, project = "seuratobject")

pattern <- paste(datasets, collapse = "|")
library(stringr)
seuratobject@meta.data$Dataset <- str_extract(colnames(seuratobject), pattern)
```

UMAP
```{r}
seuratobject <- FindVariableFeatures(seuratobject, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(seuratobject)
seuratobject <- ScaleData(seuratobject, features = all.genes)
seuratobject <- RunPCA(seuratobject, features = VariableFeatures(object = seuratobject))
ElbowPlot(seuratobject)
seuratobject <- FindNeighbors(seuratobject, dims = 1:20)
seuratobject <- FindClusters(seuratobject, resolution = 0.5)
seuratobject <- RunUMAP(seuratobject, dims = 1:10)
Idents(seuratobject) = seuratobject@meta.data$Dataset
DimPlot(seuratobject, reduction = "umap")
```

Remove batch effect using CCA 
```{r}
obj.list <- SplitObject(seuratobject, split.by = 'Dataset')
for(i in 1:length(obj.list)){
  obj.list[[i]] <- NormalizeData(object = obj.list[[i]])
  obj.list[[i]] <- FindVariableFeatures(object = obj.list[[i]])
}


features <- SelectIntegrationFeatures(object.list = obj.list)

anchors <- FindIntegrationAnchors(object.list = obj.list,
                       anchor.features = features)

seurat.integrated <- IntegrateData(anchorset = anchors)

seurat.integrated <- ScaleData(object = seurat.integrated)
seurat.integrated <- RunPCA(object = seurat.integrated)
seurat.integrated <- RunUMAP(object = seurat.integrated, dims = 1:50)

Idents(seurat.integrated) = seurat.integrated@meta.data$Dataset
DimPlot(seurat.integrated, reduction = "umap")
```

Save integrated scRNAseq object 
```{r}
saveRDS(seurat.integrated, "scRNAseq_integrated.rds")
```

Checkpoint 2

Read RDS file
```{r}
library(Seurat)
seuratobject = readRDS("scRNAseq_integrated.rds")
Idents(seuratobject) = seuratobject@meta.data$cell.annot
```

UMAP
```{r}
DefaultAssay(seuratobject) = "RNA"
Idents(seuratobject) = seuratobject@meta.data$Dataset
DimPlot(seuratobject, reduction = "umap")
```

Metacells
```{r}
library(hdWGCNA)

DefaultAssay(seuratobject) = "RNA"
data <- SetupForWGCNA(
  seuratobject,
  gene_select = "fraction", # the gene selection approach
  fraction = 0.05, # fraction of cells that a gene needs to be expressed in order to be included
  wgcna_name = "MetaCells" # the name of the hdWGCNA experiment
)

# construct metacells  in each group
seurat_obj <- MetacellsByGroups(
  seurat_obj = data,
  min_cells = 50,
  group.by = c("cell.annot", "orig.ident"), # specify the columns in seurat_obj@meta.data to group by
  reduction = 'pca', # select the dimensionality reduction to perform KNN on
  k = 25, # nearest-neighbors parameter
  max_shared = 10, # maximum number of shared cells between two metacells
  ident.group = 'cell.annot' 
)

rm(data)
gc()

metacell_obj <- GetMetacellObject(seurat_obj)

rm(seurat_obj)
gc()
```

Subset each cluster and compute TF activity
```{r}
source("src/environment_set.R")
cell_clusters = list()
batch_ids = c()
iter = 1
net = decoupleR::get_collectri(organism = 'human', split_complexes = F)
for (i in unique(metacell_obj@meta.data$cell.annot)) {
  cluster = subset(metacell_obj, idents = i, invert = FALSE)
  batch_ids = c(batch_ids, as.character(metacell_obj@meta.data[["orig.ident"]]))
  cell_clusters[[iter]] = compute.TFs.activity(ADImpute::NormalizeTPM(as.matrix(cluster@assays$RNA@data)), universe = net)
  names(cell_clusters)[iter] = i
  iter = iter + 1
}
```

Recreate sc_counts from TFs
```{r}
cell_annotations = c()
for (i in 1:length(cell_clusters)) {
  cell = c(rep(names(cell_clusters)[i], nrow(cell_clusters[[i]])))
  cell_annotations = c(cell_annotations, cell)
}
tfs_scores = do.call(rbind, unname(cell_clusters))
tfs_scores = t(tfs_scores)
tfs_scores = abs(tfs_scores)
```

Create signature
```{r}
library(omnideconv)
set_cibersortx_credentials("marcelo.hurtado@inserm.fr", "734212f6ad77fc4eea2bdb502792f294")
model_cbsx <- omnideconv::build_model(as.matrix(tfs_scores), as.character(cell_annotations),
                                      batch_ids = as.character(batch_ids), method = "cibersortx")

# model_cbsx <- omnideconv::build_model(as.matrix(seuratobject@assays$RNA@counts), as.character(seuratobject@meta.data$cell.annot), 
#                                       batch_ids = as.character(seuratobject@meta.data$orig.ident), method = "cibersortx")

saveRDS(model_cbsx, "Reference_cbsx.rds")
```

Pseudobulk
```{r}
model_cbsx = readRDS("Reference_cbsx_TF.rds")
seuratobject = readRDS("scRNAseq_integrated.rds")
DefaultAssay(seuratobject) = "RNA"
Idents(seuratobject) = seuratobject@meta.data$cell.annot

library(SingleCellExperiment)
library(glmGamPoi)
library(dplyr)
sce = as.SingleCellExperiment(seuratobject)
##Aggregating counts
aggr_counts <- glmGamPoi::pseudobulk(sce, group_by = vars(orig.ident), aggregation_functions = list(counts = "rowSums2", .default = "rowMeans2")) 
pseudo_counts = data.frame(aggr_counts@assays@data$counts) 
pseudo_counts = ADImpute::NormalizeTPM(pseudo_counts) %>%
  data.frame()

tfs_pseudo = compute.TFs.activity(pseudo_counts, universe = net)
tfs_pseudo = t(tfs_pseudo)
tfs_pseudo = abs(tfs_pseudo)

deconv_pseudo = deconvolute_cibersortx(bulk_gene_expression = tfs_pseudo, 
                                    signature = model_cbsx, verbose = T)
```

GROUNDTRUTH

Cell proportions 
```{r}
Idents(seuratobject) = seuratobject@meta.data$cell.annot
table(Idents(seuratobject)) # how many cells in each cluster
table(seuratobject$orig.ident) # how many cells per patient
table(Idents(seuratobject), seuratobject$orig.ident) # how many cells per patient in each cluster
DataClustTable = table(Idents(seuratobject), seuratobject$orig.ident) #Create data frame with number of cells per patient in each cluster
DataClustPercPat = data.frame(prop.table(table(Idents(seuratobject), seuratobject$orig.ident), margin = 2)) #Create data frame with proportion of cells per patient in each cluster
colnames(DataClustPercPat) = c("Cell_type", "Sample", "Proportion")
cells_groundtruth = DataClustPercPat

#Collapse groundtruth
library(dplyr)
cells_groundtruth <- cells_groundtruth %>%
  tidyr::pivot_wider(
    names_from = Cell_type,  
    values_from = Proportion 
  ) %>%
  tibble::column_to_rownames("Sample")
```

Groundtruth vs Deconvolution 
```{r}
cells_groundtruth <- cells_groundtruth[order(rownames(cells_groundtruth)), , drop = FALSE]
deconv_pseudo <- deconv_pseudo[order(rownames(deconv_pseudo)), , drop = FALSE]
```

Correlation plots
```{r}
library(ggplot2)
for (cell in colnames(deconv_pseudo)) {
  data = data.frame("Deconvolution" = deconv_pseudo[,cell], 
                    "Groundtruth" = cells_groundtruth[,cell])
  # Calculate correlation
  cor_test <- cor.test(data[,"Deconvolution"], data[,"Groundtruth"], method = "spearman")
    
  # Extract r and p-value
  r_value <- round(cor_test$estimate, 2)
  p_value <- signif(cor_test$p.value, 2)
    
  # Create regression plot with ggplot2
  p = ggplot(data, aes_string(x = "Deconvolution", y = "Groundtruth")) +
      geom_point() +  # Scatter plot
      geom_smooth(method = "lm", color = "blue", se = FALSE) +  # Linear regression line
      labs(x = "Deconvolution", y = "Groundtruth", title = paste0("Estimation of ", cell)) +
      annotate("text", x = Inf, y = Inf, hjust = 1.1, vjust = 2.5,
               label = paste("r =", r_value, "\np =", p_value),
               size = 4, color = "black") +
      theme_minimal()
  
  print(p)
}
```

Heatmap correlation
```{r}
cor_matrix <- cor(deconv_pseudo, cells_groundtruth)
library(pheatmap)
pheatmap(cor_matrix,
         cluster_rows = TRUE,  # Cluster rows
         cluster_cols = TRUE,  # Cluster columns
         display_numbers = TRUE, # Show correlation values
         color = colorRampPalette(c("blue", "white", "red"))(50), # Define color scale
         main = "Correlation Heatmap")
```

Distance
```{r}
dist_matrix <- 1- cor_matrix

pheatmap(dist_matrix,
         cluster_rows = TRUE,  # Cluster rows
         cluster_cols = TRUE,  # Cluster columns
         display_numbers = TRUE, # Show distance values
         color = colorRampPalette(c("white", "orange", "darkred"))(50), # Define color scale
         main = "Distance Heatmap")
```

Deconvolve mixture
```{r}
bulk = mixes$mix_rna
tfs_bulk = compute.TFs.activity(bulk) 
#tfs_bulk = tfs_bulk + abs(min(tfs_bulk))
tfs_bulk = t(tfs_bulk)
tfs_bulk = ADImpute::NormalizeTPM(tfs_bulk)
cibersortx = deconvolute_cibersortx(bulk_gene_expression = as.matrix(tfs_bulk), 
                                    signature = model_cbsx, verbose = T)
```

